---
title: "Tidying `dadmom`"
author: "Yihan Wang"
output: md_document
---

## Get data

```{r}
# don't modify this chunk 

library(tidyverse)
library(rcis)

data("dadmom")
dadmom

# hints: 
## we tested code that uses a combination of pivot_longer() and pivot_wider(); or a combination of separate(), pivot_longer(), and unite(); or only pivot_longer(). Is does not mean these are the only options to solve this question, but something to consider.
## if you use pivot_longer() and you supply multiple names to names_to, you must also supply a names_sep. Otherwise break the operation into multiple steps. If you use pivot_longer() and you get an error related to your datatype (e.g. can't combine character and double) use values_transform = as.character to simplify everything to character columns.
```

## Explain

Add a few sentences here to explain why the given data frame is not tidy. Briefly describe the approach you take to tidy it.

Explain: The original dataset has multiple observations per row, containing dad and mom, along with their incomes in one row. However, a tidy dataset should one have one observation per row, it should only contains one single parent's data (familyid, his/her parent role, name, and income).


## Tidy data

```{r}
# write here your code to tidy the data, store your tidied data as dadmom_tidy
dadmom_tidy <- 
  pivot_longer(
    data = dadmom, 
    cols = -famid, 
    names_to = c(".value", "parent"), 
    names_pattern = "(name|inc)(.)"
  )
  
# print the contents
dadmom_tidy
```


## Plot data

```{r}
# DO NOT modify this code chunk: just run it. 
# if you tided the data correctly in the previous step, it should just work. 
# parent: m or d
# inc: individual income
# famid: family ID

ggplot(data = dadmom_tidy, mapping = aes(x = parent, y = inc)) +
  geom_point() +
  geom_line(mapping = aes(group = famid)) +
  scale_y_continuous(labels = scales::dollar) +
  labs(
    title = "Gender parity and household income",
    subtitle = "Each line identifies a distinct family",
    x = "Mom or Dad",
    y = "Income",
  ) +
  theme_minimal()
```


## Session info

```{r}
# don't modify this code chunk
sessioninfo::session_info()
```

